\documentclass{article}

\begin{document}

\title{Intermediate Micro HW 2}

\maketitle

\section*{DUE DATE: June 13, 2016 at start of class}

\section{Leontief \& Substitution}

An individual has Leontief preferences over goods $x_1$ and $x_2$. He starts with income $y$ and the two goods have respective prices $p_1$ and $p_2$.

The price of good $x_2$ increases to $p_2^{'}$. Decompose this individual's change in demand into income and substitution effects.

\section{CES Demand}

Another commonly used class of utility functions is CES utility, which stands for \textbf{constant elasticity of substitution}. In general they take the form:

\[
u(x_1, x_2) = \left(\theta x_1^{\rho} + (1-\theta) x_2^{\rho}\right)^{\frac1{\rho}}
\]

If we take $\theta = \frac12$, these are equivalent to

\[
u(x_1, x_2) = \left(x_1^{\rho} + x_2^{\rho}\right)^{\frac1{\rho}}
\]

\begin{enumerate}
\item Derive the demand for CES utility with $\theta = \frac12$, i.e., find $x_1(p_1, p_2, y)$ and $x_2(p_1, p_2, y)$.
\item Decompose a change in the price of good 1 from $p_1$ to $p_1^{'}$ into substitution and income effects.
\end{enumerate}

\section{Home Production}

Bunter consumes two goods in quantities $x_1$ and $x_2$. Good 2 is a composite consumption good, and  has price 1 per unit. Consumption of$x_2$ units of good 2 requires $tx_2$ units of time, so that $t $ is the time cost per unit of good $x_2$. For example, time must be spent preparing food in order to consume it. Think of $x_1$ as {\em leisure} time that is not spent working and not spent fulflling the time-cost of consuming $x_2$. Bunter has a total of time $T$ available, and earns $w$ per unit of time spent working. All of time $T$ is consumed either in $x_1$, working, or fulfilling the time cost of consuming $x_2$. Bunter's utility function is $u(x_1, x_2) = x_1x_2^3$.

\begin{enumerate}
\item Solve Bunter's utility maximization problem to find his demand functions for goods 1 and 2.

\item  Suppose $t$ decreases. For example, new technologies maydecrease food preparation time. What is the effect on the consumptionof good 2? What is the effect on the amount of time this person spendsworking? Be precise.

\item  Suppose the wage $w$ increases. What is the effect onthe amount of time this person spends working, and on the amount ofleisure time, $x_1$ she consumes?
\end{enumerate}

\end{document}